---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
activity <- read.csv(file = 'activity.csv', stringsAsFactors = FALSE)

require(magrittr)
require(dplyr)
```

## What is mean total number of steps taken per day?

The prompt says that we can `ignore missing values`.  I am taking that to mean that we can `na.rm` when summing by day.

### steps per day
```{r}

per_day <- activity %>%
  dplyr::group_by(date) %>%
  dplyr::summarize(
    total_steps = sum(steps, na.rm = TRUE)
  ) 

per_day

```

### histogram
```{r}
per_day$total_steps %>% 
  hist(
    breaks = seq(0, 25000, 2500), 
    main = 'Steps per Day'
  )
```

### mean
```{r}
per_day$total_steps %>% mean()
```

### median
```{r}
per_day$total_steps %>% median()
```

## What is the average daily activity pattern?

```{r}

per_interval <- activity %>%
  dplyr::group_by(interval) %>%
  dplyr::summarize(
    total_steps = mean(steps, na.rm = TRUE)
  )

per_interval
```

```{r}
plot(
  per_interval$interval, per_interval$total_steps, type='l',
  main = 'Average Steps per 5 min Daily Interval', 
  xlab = 'Interval', ylab = 'Average Steps', 
  col = 'hotpink', lwd = 2,
)

```

### largest average interval
The interval with the largest (maximum) number of steps is interval 835, with 206.2 steps.
```{r}
per_interval %>% arrange(-total_steps)
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
