---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
activity <- read.csv(file = 'activity.csv', stringsAsFactors = FALSE)

require(magrittr)
require(dplyr)
```

## What is mean total number of steps taken per day?

The prompt says that we can `ignore missing values`.  I am taking that to mean that we can `na.rm` when summing by day.

### steps per day
```{r}

per_day <- activity %>%
  dplyr::group_by(date) %>%
  dplyr::summarize(
    avg_steps = sum(steps, na.rm = TRUE)
  ) 

per_day

```

### histogram
```{r}
per_day$avg_steps %>% 
  hist(
    breaks = seq(0, 25000, 2500), 
    main = 'Steps per Day'
  )
```

### mean
```{r}
per_day$avg_steps %>% mean()
```

### median
```{r}
per_day$avg_steps %>% median()
```

## What is the average daily activity pattern?

```{r}

per_interval <- activity %>%
  dplyr::group_by(interval) %>%
  dplyr::summarize(
    avg_steps = mean(steps, na.rm = TRUE)
  )

per_interval
```

```{r}
plot(
  per_interval$interval, per_interval$avg_steps, type='l',
  main = 'Average Steps per 5 min Daily Interval', 
  xlab = 'Interval', ylab = 'Average Steps', 
  col = 'hotpink', lwd = 2,
)

```

### largest average interval
The interval with the largest (maximum) number of steps is interval **835**, with 206.2 steps.
```{r}
per_interval %>% arrange(-avg_steps)
```

## Imputing missing values

### total number of missing values
There are 2304 missing (NA) rows.

```{r}
activity$steps %>% is.na() %>% sum()
```

### devise and create an imputation strategy

```{r}

ma <- function(x, n = 5){
  stats::filter(x, rep(1/n, n), sides = 1)
}

per_day$rolling_avg <- ma(per_day$avg_steps, 3)
per_day$rolling_avg <- ifelse(per_day$rolling_avg %>% is.na(), 0, per_day$rolling_avg)
per_day <- per_day %>%
  ungroup() %>%
  rowwise() %>%
  mutate(
    avg_steps = avg_steps %>% as.numeric(),
    imputed_steps = ifelse(
      avg_steps == 0, rolling_avg, avg_steps
    )
  )

per_day

```

### histogram
```{r}
per_day$imputed_steps %>% 
  hist(
    breaks = seq(0, 25000, 2500), 
    main = 'Steps per Day'
  )
```

### mean
```{r}
per_day$imputed_steps %>% mean()
```

### median
```{r}
per_day$imputed_steps %>% median()
```

### discussion
My strategy for dealing with missing data is not very agressive - my general theory is that days with all NAs/missing data are almost certainly a mistake in the data (having owned a fitbit, those are probably days you forgot to wear it!).

I computed a 3 day trailing average, and on those days with no steps, I replace the missing days with the trailing average.  This mainly impacts the lower tail of the distribution - rather than having some days that are zero, those days now look more like their neighbors, in terms of step count.

This brings the mean up fairly significantly, from ~9,350 to ~10,050. The median, however, is not changed - my imputed data set still has a median step count of 10,395.  I think that's because all of the days that I imputed were already below the median, and while they moved up, no days moved up _past_ the median (and no days were added) meaning that the median value remained unchanged.

## Are there differences in activity patterns between weekdays and weekends?

### create new variable
```{r}

by_weekdays <- activity %>%
  mutate(
    is_weekday = weekdays(date %>% as.Date()),
    is_weekday = ! is_weekday %in% c('Saturday', 'Sunday'),
    is_weekday = ifelse(is_weekday, 'Weekdays', 'Weekends')
  ) %>%
  group_by(is_weekday, interval) %>%
  dplyr::summarize(
    avg_steps = mean(steps, na.rm = TRUE),
    n = n()
  )

by_weekdays
```

### make a panel plot
```{r}
require(ggplot2)

ggplot(
  data = by_weekdays,
  aes(
    x = interval,
    y = avg_steps,
    color = is_weekday
  )
) +
geom_line() +
theme_bw() +
facet_grid(is_weekday ~ .)


```

Activity starts earlier in the day on weekdays, and is more concentrated in the earlier intervals.  Weekend activity is much more frequent between intervals 1000 and 1500.